---
title: "anglerDiaryChoice"
output: html_document
date: "2024-01-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, lubridate, here, apollo)
```

Read in angler diary data, add column specifying inland lake, river, or great lakes
Classifying impoundments/riverine lakes as inland lakes. 

Sets of choice models: 
- Participation (each day of month--fish or no fish)
- For those participating: Waterbody type (inland lake, great lake, river)
- For those fishing inland lakes: target species (walleye, bluegill, something else)
- For those fishing walleye and/or bluegill: fishing site
This last one will need some iterating. Try: 
- as cross nested model: Distance and species presence (identifiable inland lakes--diary data)
  - Compare to DCE estimates: only distance and species presence (non-zero expected catch rate)
- separate models for first fishing trip recorded and later trips
- latent classes (no covariates?)


"Coastal" counties are counties touching, Lake Superior, Lake Michigan, or the Mississippi or Wisconsin rivers. (Rivers (sort of) account for the Driftless area, which has few inland lakes)

```{r}
`%!in%`<-Negate(`%in%`)

codes<-read_csv(here::here("data","coding.csv"))%>%
  select(code, `Target Species`, category)%>%
  rename("TargetSpecies"=code,
         "TargetSpeciesName"=`Target Species`, 
         "TargetSpeciesCat"=category)

coastal<-read_csv(here::here("data", "coastalCounties.csv"))

anglerInfo<-read_csv(here::here("data","angler_info.csv"))


diaryDataWaterbody<-read_csv(here::here("data","angler.diary.data.csv"))%>%
    arrange(SurveyID)%>%
  mutate(daysMonth=ifelse(Month%in%c(9,4,6,11), 30, 
                          ifelse(Month==2, 28, 31)),
         date_month=as.numeric(month(mdy(Date))),
         outsideMonth=ifelse(FishedYN==1 & date_month!=Month,1,0),
         Year=ifelse(Month%in%c(4,5,6,7,8,9,10,11,12), 2021, 2022))%>%
  # remove entries where participants went fishing outside the sampled calendar month
  filter(outsideMonth==0)%>%
  filter(Mode==1)%>%
  filter(SurveyID!="999999")%>%
  filter(SurveyID%!in%c(1:10))%>%
  group_by(SurveyID, Date)%>%
  # for now, removing second or third trips within the same day and angler. Relatively low sample (229) and would require re-working model approach. 
  # still, missed opportunity to capture 'lake surfing'. Think about it some more. One angler went to 5 different lakes!
  mutate(lakeNum=row_number())%>%
  # removing additional lakes visited in 1 day
  filter(lakeNum%!in%c(2:5))%>%
  ungroup()%>%
  left_join(codes, by="TargetSpecies")%>%
  mutate(waterbodyType=ifelse(grepl("River",WaterFished), "river",
                              ifelse(WaterFished%in%c("Lake Michigan", "Lake Superior"), "greatLake",
                              "inlandLake")))%>%
  left_join(anglerInfo, by="SurveyID")%>%
  mutate(coastOrRiver=ifelse(PrimaryCounty%in%coastal$county, 1, 0))
  # this is an ungainly number of columns. Cutting out catch and keep numbers for now. Can join it back by surveyID later
  
catchKeep<-diaryDataWaterbody%>%
  select(SurveyID, BluegillNumCaught:OtherNumKeep)

diaryData<-diaryDataWaterbody%>%
  select(!BluegillNumCaught:OtherNumKeep)

```

Now start filtering out entries that don't have the county of the waterFished

```{r}
noCounty<-diaryData%>%
  mutate(noCounty=ifelse(County=="99" | WaterFished=="99",1,0))%>%
  group_by(SurveyID)%>%
  summarize(nNoCounty=sum(noCounty))%>%
  filter(nNoCounty>0)%>%
  ungroup()

diaryData.f<-filter(diaryData, SurveyID%!in%noCounty$SurveyID)%>%
  select(!c(Gender, YearOfBirth, Residency, PrimaryCounty, "Letter Version", BRM, Responded, Fished))

```

Not everyone labeled their target species, so check that before moving to next step of modeling

Check that I don't have duplicates or multiple months
11593 unique surveyIDs, neither spanning multiple months

```{r}
uniqueID<-distinct(diaryData.f[,c("SurveyID","Month","Year","daysMonth")])
# some data spreads across more than one month. That's fine I guess. But check on those
# Update: it was not, but now fixed. Survey was meant to take place over 1 calendar month. 
 check<-uniqueID%>%
   group_by(SurveyID)%>%
   summarize(nMonth=length(unique(Month)))

```

Spread the data to record angler decisions to fish (or not) each day of the diary month

```{r}

fished<-filter(diaryDataWaterbody, FishedYN==1)
noFished<-filter(diaryDataWaterbody, FishedYN==2)

dataSpread<-uniqueID%>%
  left_join(anglerInfo, by="SurveyID")%>%
    slice(rep(1:n(), times=daysMonth))%>%
  group_by(SurveyID)%>%
  mutate(day=row_number())%>%
  ungroup()%>%
  # add a date column by pasting month, day, year
  # previously messed up dates
  mutate(Date=paste(Diary_Month, day, Year, sep="/"))%>%
  # indicator for if they fished at all that month
  mutate(fishedEver=ifelse(SurveyID%in%fished$SurveyID, 1, ifelse(SurveyID%in%noFished$SurveyID, 0, "error")))%>%
  select(!c("Month","Year","Diary_Month", "daysMonth"))%>%
  left_join(diaryData.f, by=c("SurveyID", "Date"))%>%
  # FishedYN now only has 1 or NA (and 1 incorrect 2 on a day recorded fishing)
  mutate(FishedYN=ifelse(FishedYN%in%c(1,2), 1, 0))%>%
  # remove columns we don't need
  select(!c(day, ID, Mode, Month, date_month, outsideMonth, Year))

```
Next add on weather data

```{r}
precip<-read_csv(here::here("data","precip.county.date.csv"))%>%
  rename("County"=county)%>%
  mutate(Date=mdy(Date),
         year=year(Date))

#choiceData<-left_join(dataSpread, precip, by=c("County", "Date"))

# missing days?

precipCheck<-precip%>%
  group_by(County, year)%>%
  summarize(nDays=length(unique(Date)))
precipCheck
```
Ah. The work I did with precipitation data (angler diary data directory, nested.logit.wranging.Rmd) only got
precipitation for all days fished. 

```{r}
choiceData<-dataSpread

write.csv(choiceData, here::here("data","choice.data.participation.waterbody.type.csv"))
```

