---
title: "nested choice model"
output: html_document
date: "2024-01-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, here, lubridate, apollo, fastDummies)
```

This script reads in choice data from angler diaries, fits nested logit model predicting 
fishing participation (yes, no) and type of waterbody (river, great lake, inland lake)

```{r}
choiceData<-read_csv(here::here("data","choice.data.participation.waterbody.type.csv"))
choiceData<-choiceData[,-1]

holidays<-mdy(c("05/31/2021","09/06/2021","11/25/2021","12/31/2021",
                "01/17/2022"))

coastal<-read_csv(here::here("data","coastalCounties.csv"))

river<-coastal%>%
  filter(grepl("River", waterbody))

greatLake<-coastal%>%
  filter(grepl("Lake", waterbody))

choiceData.cov<-choiceData%>%
  mutate(Date=mdy(Date),
         month=as.character(month(Date, label=T)),
         weekday=wday(Date, label=T),
         weekendHoliday=ifelse(weekday%in%c("Sun","Sat") | Date%in%holidays, 1, 0),
         coastal_greatLake=ifelse(PrimaryCounty%in%greatLake$county, 1, 0),
         coastal_river=ifelse(PrimaryCounty%in%river$county, 1, 0),
         age=2022-YearOfBirth)%>%
  select(SurveyID, age, PrimaryCounty, Date, fishedEver, FishedYN, WaterFished, County, Hours:waterbodyType,
         coastal_greatLake, coastal_river, month, weekday, weekendHoliday)%>%
  mutate(age.sc=scale(age, center=T, scale=T),
         choice.partic=ifelse(FishedYN==0, "noFish", "fish"),
         choice.waterbody=ifelse(FishedYN==0, "noFish",
                       ifelse(waterbodyType=="river", "river",
                              ifelse(waterbodyType=="greatLake", "greatLake",
                                     ifelse(waterbodyType=="inlandLake", "inlandLake","error")))),
         choice=ifelse(fishedEver==0, "neverFished",
                       ifelse(FishedYN==0, "notToday",
                              ifelse(waterbodyType=="river","river",
                                     ifelse(waterbodyType=="greatLake", "greatLake",
                                            ifelse(waterbodyType=="inlandLake", "inlandLake", "error"))))),
         age_weekend=age.sc*weekendHoliday)%>%
  mutate(shoulder=ifelse(month%in%c("Oct","Nov","Mar","Apr"), 1, 0),
         peakSummer=ifelse(month%in%c("May","Jun","Jul"), 1, 0))



choiceData.cov.f<-choiceData%>%
  mutate(Date=mdy(Date),
         month=as.character(month(Date, label=T)),
         weekday=wday(Date, label=T),
         weekendHoliday=ifelse(weekday%in%c("Sun","Sat") | Date%in%holidays, 1, 0),
         coastal_greatLake=ifelse(PrimaryCounty%in%greatLake$county, 1, 0),
         coastal_river=ifelse(PrimaryCounty%in%river$county, 1, 0),
         age=2022-YearOfBirth)%>%
  select(SurveyID, age, PrimaryCounty, Date, fishedEver, FishedYN, WaterFished, County, Hours:waterbodyType,
         coastal_greatLake, coastal_river, month, weekday, weekendHoliday)%>%
  mutate(age.sc=scale(age, center=T, scale=T),
         choice.partic=ifelse(FishedYN==0, "noFish", "fish"),
         choice.waterbody=ifelse(FishedYN==0, "noFish",
                       ifelse(waterbodyType=="river", "river",
                              ifelse(waterbodyType=="greatLake", "greatLake",
                                     ifelse(waterbodyType=="inlandLake", "inlandLake","error")))),
         age_weekend=age.sc*weekendHoliday)%>%
  filter(fishedEver==1)%>%
  mutate(shoulder=ifelse(month%in%c("Oct","Nov","Mar","Apr"), 1, 0),
         peakSummer=ifelse(month%in%c("May","Jun","Jul"), 1, 0))

choiceData.dum<-dummy_cols(choiceData.cov, select_columns="month",
                           remove_first_dummy=F,
                           remove_selected_columns=T)


choiceData.dum.f<-dummy_cols(choiceData.cov.f, select_columns="month",
                           remove_first_dummy=F,
                           remove_selected_columns=T)
```
Jumping straight to  nested logit didn't work, shocking no one but myself. 

Start simpler with logit: fishing vs not

```{r}

### Initialise code
apollo_initialise()

### Set core controls
apollo_control = list(
  modelName       = "logit_participation",
  modelDescr      = "Simple logit model--participation",
  indivID         = "SurveyID", 
  outputDirectory = "output_logit_participation"
)

# ################################################################# #
#### LOAD DATA AND APPLY ANY TRANSFORMATIONS                     ####
# ################################################################# #

### Loading data from package
### if data is to be loaded from a file (e.g. called data.csv), 
### the code would be: database = read.csv("data.csv",header=TRUE)
database = choiceData.dum
### for data dictionary, use ?apollo_modeChoiceData


# ################################################################# #
#### DEFINE MODEL PARAMETERS                                     ####
# ################################################################# #

### Vector of parameters, including any that are kept fixed in estimation
apollo_beta=c(asc_noFish=1,
              asc_fish=0,
               b_weekend_holiday=0,
              # b_shoulder=0,
              # b_peakSummer=0,
               b_age=0,
               b_age_weekend = 0,
                b_month_Feb=0,
                b_month_Mar=0,
                b_month_Apr=0,
                b_month_May=0,
                b_month_Jun=0,
                b_month_Jul=0,
                b_month_Aug=0,
                b_month_Sep=0,
                b_month_Oct=0,
                # no November data
                b_month_Dec=0
              )

### Vector with names (in quotes) of parameters to be kept fixed at their starting value in apollo_beta, use apollo_beta_fixed = c() if none
apollo_fixed = c("asc_noFish")

# ################################################################# #
#### GROUP AND VALIDATE INPUTS                                   ####
# ################################################################# #

apollo_inputs = apollo_validateInputs()

# ################################################################# #
#### DEFINE MODEL AND LIKELIHOOD FUNCTION                        ####
# ################################################################# #

apollo_probabilities=function(apollo_beta, apollo_inputs, functionality="estimate"){

  ### Attach inputs and detach after function exit
  apollo_attach(apollo_beta, apollo_inputs)
  on.exit(apollo_detach(apollo_beta, apollo_inputs))

  ### Create list of probabilities P
  P = list()
  
  ### List of utilities: these must use the same names as in mnl_settings, order is irrelevant
  V = list()
  V[["noFish"]]  = asc_noFish
  V[["fish"]]  = asc_fish + b_weekend_holiday * weekendHoliday + b_month_Feb * month_Feb + 
    b_month_Mar * month_Mar + b_month_Apr * month_Apr + b_month_May * month_May +
    b_month_Jun * month_Jun + b_month_Jul * month_Jul + b_month_Aug * month_Aug +
    b_month_Sep * month_Sep + b_month_Oct * month_Oct + b_month_Dec * month_Dec +
    b_age * age.sc + b_age_weekend * age_weekend


  
  ### Define settings for MNL model component
  mnl_settings = list(
    alternatives  = c(noFish="noFish", fish="fish"), 
    choiceVar     = choice.partic,
    utilities     = V
  )
  
  ### Compute probabilities using MNL model
  P[["model"]] = apollo_mnl(mnl_settings, functionality)

  ### Take product across observation for same individual
  P = apollo_panelProd(P, apollo_inputs, functionality)

  ### Prepare and return outputs of function
  P = apollo_prepareProb(P, apollo_inputs, functionality)
  return(P)
}

# ################################################################# #
#### MODEL ESTIMATION                                            ####
# ################################################################# #

model = apollo_estimate(apollo_beta, apollo_fixed, apollo_probabilities, apollo_inputs)

# ################################################################# #
#### MODEL OUTPUTS                                               ####
# ################################################################# #

# ----------------------------------------------------------------- #
#---- FORMATTED OUTPUT (TO SCREEN)                               ----
# ----------------------------------------------------------------- #

apollo_modelOutput(model)

# ----------------------------------------------------------------- #
#---- FORMATTED OUTPUT (TO FILE, using model name)               ----
# ----------------------------------------------------------------- #

apollo_saveOutput(model)


```

Instead of nested logit (tried below--problems with nesting parameter), trying normal MNL with covariates.
If that works out, try MMNL

```{r}
# ################################################################# #
#### LOAD LIBRARY AND DEFINE CORE SETTINGS                       ####
# ################################################################# #


### Initialise code
apollo_initialise()

### Set core controls
apollo_control = list(
  modelName       = "MNL_participate_waterbodyType",
  modelDescr      = "MNL participation and waterbody type",
  indivID         = "SurveyID", 
  outputDirectory = "output_mnl_fish_waterbody"
)

# ################################################################# #
#### LOAD DATA AND APPLY ANY TRANSFORMATIONS                     ####
# ################################################################# #

### Loading data from package
### if data is to be loaded from a file (e.g. called data.csv), 
### the code would be: database = read.csv("data.csv",header=TRUE)
database = choiceData.dum
### for data dictionary, use ?apollo_modeChoiceData


# ################################################################# #
#### DEFINE MODEL PARAMETERS                                     ####
# ################################################################# #

### Vector of parameters, including any that are kept fixed in estimation. no observations in november, left out on purpose. 
apollo_beta=c( asc_noFish=0,
               asc_fish=0,
              b_weekend_holiday=0,
              b_age=0,
              b_weekend_age=0,
                b_month_Feb=0,
                b_month_Mar=0,
                b_month_Apr=0,
                b_month_May=0,
                b_month_Jun=0,
                b_month_Jul=0,
                b_month_Aug=0,
                b_month_Sep=0,
                b_month_Oct=0,
                # no November data
                b_month_Dec=0,
              b_coastal_greatLake=0,
              b_coastal_river=0

              )

### Vector with names (in quotes) of parameters to be kept fixed at their starting value in apollo_beta, use apollo_beta_fixed = c() if none
apollo_fixed = c("asc_noFish")

# ################################################################# #
#### GROUP AND VALIDATE INPUTS                                   ####
# ################################################################# #

apollo_inputs = apollo_validateInputs()

# ################################################################# #
#### DEFINE MODEL AND LIKELIHOOD FUNCTION                        ####
# ################################################################# #

apollo_probabilities=function(apollo_beta, apollo_inputs, functionality="estimate"){

  ### Attach inputs and detach after function exit
  apollo_attach(apollo_beta, apollo_inputs)
  on.exit(apollo_detach(apollo_beta, apollo_inputs))

  ### Create list of probabilities P
  P = list()
  
  ### create alternative specific constants
  
  asc_fish_value = asc_fish + b_weekend_holiday * weekendHoliday + b_month_Feb * month_Feb + 
    b_month_Mar * month_Mar + b_month_Apr * month_Apr + b_month_May * month_May +
    b_month_Jun * month_Jun + b_month_Jul * month_Jul + b_month_Aug * month_Aug +
    b_month_Sep * month_Sep + b_month_Oct * month_Oct + b_month_Dec * month_Dec +
    b_age * age.sc + b_weekend_age * age_weekend
  
  ### List of utilities: these must use the same names as in mnl_settings, order is irrelevant

  V = list()
  V[["noFish"]] = asc_noFish 
  V[["river"]] = asc_fish_value + b_coastal_river * coastal_river
  V[["greatLake"]] = asc_fish_value + b_coastal_greatLake * coastal_greatLake
  V[["inlandLake"]] = asc_fish_value

 ### Define settings for MNL model component
  mnl_settings = list(
    alternatives = c(noFish="noFish",river="river",greatLake="greatLake",inlandLake="inlandLake"),
    choiceVar    = choice.waterbody,
    utilities    = V
  )
  
  ### Compute probabilities using MNL model
  P[["model"]] = apollo_mnl(mnl_settings, functionality)
  
  ### Take product across observation for same individual
  P = apollo_panelProd(P, apollo_inputs, functionality)
  
  ### Prepare and return outputs of function
  P = apollo_prepareProb(P, apollo_inputs, functionality)
  return(P)
}

# ################################################################# #
#### MODEL ESTIMATION                                            ####
# ################################################################# #

model = apollo_estimate(apollo_beta, apollo_fixed, apollo_probabilities, apollo_inputs)

# ################################################################# #
#### MODEL OUTPUTS                                               ####
# ################################################################# #

# ----------------------------------------------------------------- #
#---- FORMATTED OUTPUT (TO SCREEN)                               ----
# ----------------------------------------------------------------- #

apollo_modelOutput(model)

# ----------------------------------------------------------------- #
#---- FORMATTED OUTPUT (TO FILE, using model name)               ----
# ----------------------------------------------------------------- #

apollo_saveOutput(model)

# ################################################################# #
##### POST-PROCESSING                                            ####
# ################################################################# #

### Print outputs of additional diagnostics to new output file (remember to close file writing when complete)
apollo_sink()
```





converged! But still has the problem of a negative lambda value, which isn't allowed. Maybe this nesting structure isn't appropriate? 

```{r}
# ################################################################# #
#### LOAD LIBRARY AND DEFINE CORE SETTINGS                       ####
# ################################################################# #


### Initialise code
apollo_initialise()

### Set core controls
apollo_control = list(
  modelName       = "NL_participate_waterbodyType",
  modelDescr      = "Nested logit participation and waterbody type",
  indivID         = "SurveyID", 
  outputDirectory = "output_nested_fish_waterbody"
)

# ################################################################# #
#### LOAD DATA AND APPLY ANY TRANSFORMATIONS                     ####
# ################################################################# #

### Loading data from package
### if data is to be loaded from a file (e.g. called data.csv), 
### the code would be: database = read.csv("data.csv",header=TRUE)
database = choiceData.dum
### for data dictionary, use ?apollo_modeChoiceData


# ################################################################# #
#### DEFINE MODEL PARAMETERS                                     ####
# ################################################################# #

### Vector of parameters, including any that are kept fixed in estimation. no observations in november, left out on purpose. 
apollo_beta=c( asc_noFish=1,
              asc_river=0,
              asc_greatLake=0,
              asc_inland=0,
              b_weekend_holiday=0,
              #b_age=0,
              #b_weekend_age=0,
                b_month_Feb=0,
                b_month_Mar=0,
                b_month_Apr=0,
                b_month_May=0,
                b_month_Jun=0,
                b_month_Jul=0,
                b_month_Aug=0,
                b_month_Sep=0,
                b_month_Oct=0,
                # no November data
                b_month_Dec=0,
              b_coastal_greatLake=0,
              b_coastal_river=0,
              lambda_fish=1
              
              )

### Vector with names (in quotes) of parameters to be kept fixed at their starting value in apollo_beta, use apollo_beta_fixed = c() if none
apollo_fixed = c("asc_noFish")

# ################################################################# #
#### GROUP AND VALIDATE INPUTS                                   ####
# ################################################################# #

apollo_inputs = apollo_validateInputs()

# ################################################################# #
#### DEFINE MODEL AND LIKELIHOOD FUNCTION                        ####
# ################################################################# #

apollo_probabilities=function(apollo_beta, apollo_inputs, functionality="estimate"){

  ### Attach inputs and detach after function exit
  apollo_attach(apollo_beta, apollo_inputs)
  on.exit(apollo_detach(apollo_beta, apollo_inputs))

  ### Create list of probabilities P
  P = list()
  
  ### create alternative specific constants
  
  asc_river_value = asc_river + b_coastal_river * coastal_river 
  asc_greatLake_value = asc_greatLake + b_coastal_greatLake * coastal_greatLake 

  
  ### List of utilities: these must use the same names as in mnl_settings, order is irrelevant

  V = list()
  V[["noFish"]] = asc_noFish 
  V[["river"]] = asc_river_value + b_weekend_holiday * weekendHoliday + b_month_Feb * month_Feb + 
    b_month_Mar * month_Mar + b_month_Apr * month_Apr + b_month_May * month_May +
    b_month_Jun * month_Jun + b_month_Jul * month_Jul + b_month_Aug * month_Aug +
    b_month_Sep * month_Sep + b_month_Oct * month_Oct + b_month_Dec * month_Dec
  V[["greatLake"]] = asc_greatLake_value + b_weekend_holiday * weekendHoliday + b_month_Feb * month_Feb + 
    b_month_Mar * month_Mar + b_month_Apr * month_Apr + b_month_May * month_May +
    b_month_Jun * month_Jun + b_month_Jul * month_Jul + b_month_Aug * month_Aug +
    b_month_Sep * month_Sep + b_month_Oct * month_Oct + b_month_Dec * month_Dec
  V[["inlandLake"]] = asc_inland + b_weekend_holiday * weekendHoliday + b_month_Feb * month_Feb + 
    b_month_Mar * month_Mar + b_month_Apr * month_Apr + b_month_May * month_May +
    b_month_Jun * month_Jun + b_month_Jul * month_Jul + b_month_Aug * month_Aug +
    b_month_Sep * month_Sep + b_month_Oct * month_Oct + b_month_Dec * month_Dec

  # Specify  nests for NL model
  
  nlNests = list(root = 1,
                 fish = lambda_fish)
  
  # specify tree structure for NL model
  nlStructure = list()
  nlStructure[["root"]] = c("noFish","fish")
  nlStructure[["fish"]] = c("river","greatLake","inlandLake")
  
  nl_settings <-list(
    alternatives = c(noFish="noFish", river="river", inlandLake="inlandLake", greatLake="greatLake"),
    utilities=V,
    choiceVar = choice.waterbody,
    nlNests = nlNests,
    nlStructure = nlStructure
  )

 ### Compute probabilities using NL model
  P[["model"]] = apollo_nl(nl_settings, functionality)
  
  ### Take product across observation for same individual
  P = apollo_panelProd(P, apollo_inputs, functionality)
  
  ### Prepare and return outputs of function
  P = apollo_prepareProb(P, apollo_inputs, functionality)
  return(P)
}
# ################################################################# #
#### MODEL ESTIMATION                                            ####
# ################################################################# #

model = apollo_estimate(apollo_beta, apollo_fixed, apollo_probabilities, apollo_inputs)

# ################################################################# #
#### MODEL OUTPUTS                                               ####
# ################################################################# #

# ----------------------------------------------------------------- #
#---- FORMATTED OUTPUT (TO SCREEN)                               ----
# ----------------------------------------------------------------- #

apollo_modelOutput(model)

# ----------------------------------------------------------------- #
#---- FORMATTED OUTPUT (TO FILE, using model name)               ----
# ----------------------------------------------------------------- #

saveOutput_settings=list(
  printPVal=2
)


apollo_saveOutput(model, saveOutput_settings)


```
What if.. three level nested logit: 
- Never fish
- fish sometime
  - not today
  - fish today
    - inland
    - great lake
    - river
    
    
lambda values are still super weird. try constraining them
```{r}
# ################################################################# #
#### LOAD LIBRARY AND DEFINE CORE SETTINGS                       ####
# ################################################################# #


### Initialise code
apollo_initialise()

### Set core controls
apollo_control = list(
  modelName       = "NL_3_level",
  modelDescr      = "Nested logit participation and waterbody type--3 level",
  indivID         = "SurveyID", 
  outputDirectory = "output_nested_3_level"
)

# ################################################################# #
#### LOAD DATA AND APPLY ANY TRANSFORMATIONS                     ####
# ################################################################# #

### Loading data from package
### if data is to be loaded from a file (e.g. called data.csv), 
### the code would be: database = read.csv("data.csv",header=TRUE)
database = choiceData.dum
### for data dictionary, use ?apollo_modeChoiceData


# ################################################################# #
#### DEFINE MODEL PARAMETERS                                     ####
# ################################################################# #

### Vector of parameters, including any that are kept fixed in estimation. no observations in november, left out on purpose. 
apollo_beta=c( asc_neverFished=1,
               asc_notToday=1,
              asc_river=0,
              asc_greatLake=0,
              asc_inland=0,
              b_weekend_holiday=0,
              #b_age=0,
              #b_weekend_age=0,
                b_month_Feb=0,
                b_month_Mar=0,
                b_month_Apr=0,
                b_month_May=0,
                b_month_Jun=0,
                b_month_Jul=0,
                b_month_Aug=0,
                b_month_Sep=0,
                b_month_Oct=0,
                # no November data
                b_month_Dec=0,
              b_coastal_greatLake=0,
              b_coastal_river=0,
              lambda_fished=1,
              lambda_today=1
              
              )

### Vector with names (in quotes) of parameters to be kept fixed at their starting value in apollo_beta, use apollo_beta_fixed = c() if none
apollo_fixed = c("asc_neverFished")

# ################################################################# #
#### GROUP AND VALIDATE INPUTS                                   ####
# ################################################################# #

apollo_inputs = apollo_validateInputs()

# ################################################################# #
#### DEFINE MODEL AND LIKELIHOOD FUNCTION                        ####
# ################################################################# #

apollo_probabilities=function(apollo_beta, apollo_inputs, functionality="estimate"){

  ### Attach inputs and detach after function exit
  apollo_attach(apollo_beta, apollo_inputs)
  on.exit(apollo_detach(apollo_beta, apollo_inputs))

  ### Create list of probabilities P
  P = list()
  
  ### create alternative specific constants
  
  asc_river_value = asc_river + b_coastal_river * coastal_river 
  asc_greatLake_value = asc_greatLake + b_coastal_greatLake * coastal_greatLake 

  
  ### List of utilities: these must use the same names as in mnl_settings, order is irrelevant

  V = list()
  V[["neverFished"]] = asc_neverFished
  V[["notToday"]] = asc_notToday
  V[["river"]] = asc_river_value + b_weekend_holiday * weekendHoliday + b_month_Feb * month_Feb + 
    b_month_Mar * month_Mar + b_month_Apr * month_Apr + b_month_May * month_May +
    b_month_Jun * month_Jun + b_month_Jul * month_Jul + b_month_Aug * month_Aug +
    b_month_Sep * month_Sep + b_month_Oct * month_Oct + b_month_Dec * month_Dec
  V[["greatLake"]] = asc_greatLake_value + b_weekend_holiday * weekendHoliday + b_month_Feb * month_Feb + 
    b_month_Mar * month_Mar + b_month_Apr * month_Apr + b_month_May * month_May +
    b_month_Jun * month_Jun + b_month_Jul * month_Jul + b_month_Aug * month_Aug +
    b_month_Sep * month_Sep + b_month_Oct * month_Oct + b_month_Dec * month_Dec
  V[["inlandLake"]] = asc_inland + b_weekend_holiday * weekendHoliday + b_month_Feb * month_Feb + 
    b_month_Mar * month_Mar + b_month_Apr * month_Apr + b_month_May * month_May +
    b_month_Jun * month_Jun + b_month_Jul * month_Jul + b_month_Aug * month_Aug +
    b_month_Sep * month_Sep + b_month_Oct * month_Oct + b_month_Dec * month_Dec

  # Specify  nests for NL model
  
  nlNests = list(root = 1,
                 fished = lambda_fished,
                 today = lambda_today)
  
  # specify tree structure for NL model
  nlStructure = list()
  nlStructure[["root"]] = c("neverFished","fished")
  nlStructure[["fished"]] = c("notToday", "today")
  nlStructure[["today"]] = c("river","greatLake","inlandLake")
  
  nl_settings <-list(
    alternatives = c(neverFished="neverFished", notToday="notToday", river="river", 
                     inlandLake="inlandLake", greatLake="greatLake"),
    utilities=V,
    choiceVar = choice,
    nlNests = nlNests,
    nlStructure = nlStructure
  )

 ### Compute probabilities using NL model
  P[["model"]] = apollo_nl(nl_settings, functionality)
  
  ### Take product across observation for same individual
  P = apollo_panelProd(P, apollo_inputs, functionality)
  
  ### Prepare and return outputs of function
  P = apollo_prepareProb(P, apollo_inputs, functionality)
  return(P)
}
# ################################################################# #
#### MODEL ESTIMATION                                            ####
# ################################################################# #

model = apollo_estimate(apollo_beta, apollo_fixed, apollo_probabilities, apollo_inputs)

# ################################################################# #
#### MODEL OUTPUTS                                               ####
# ################################################################# #

# ----------------------------------------------------------------- #
#---- FORMATTED OUTPUT (TO SCREEN)                               ----
# ----------------------------------------------------------------- #

apollo_modelOutput(model)

# ----------------------------------------------------------------- #
#---- FORMATTED OUTPUT (TO FILE, using model name)               ----
# ----------------------------------------------------------------- #

saveOutput_settings=list(
  printPVal=2
)


apollo_saveOutput(model, saveOutput_settings)

apollo_sink()
```
lambda values (nesting parameters) are out of control. Constraining it didn't help and isn't advisable anyway


suggestion to instead try mixed logit for panel data https://www.statalist.org/forums/forum/general-stata-discussion/general/1569500-nested-logit-with-panel-data 
    