---
title: "nested choice model"
output: html_document
date: "2024-01-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, here, lubridate, apollo, fastDummies, RColorBrewer)
```

This script reads in choice data from angler diaries, fits nested logit model predicting 
fishing participation (yes, no) and type of waterbody (river, great lake, inland lake)

```{r}
choiceData<-read_csv(here::here("data","choice.data.participation.waterbody.type.csv"))
choiceData<-choiceData[,-1]

greatLake<-choiceData%>%
  filter(WaterFished%in%c("Lake Michigan","Lake Superior"))%>%
  group_by(WaterFished)%>%
  summarize(n=n())
# hm, not that many.

holidays<-mdy(c("05/31/2021","09/06/2021","11/25/2021","12/31/2021",
                "01/17/2022"))

coastal<-read_csv(here::here("data","coastalCounties.csv"))

river<-coastal%>%
  filter(grepl("River", waterbody))

greatLake<-coastal%>%
  filter(grepl("Lake", waterbody))

lakeMichigan<-coastal%>%
  filter(waterbody=="Lake Michigan")

lakeSuperior<-coastal%>%
  filter(waterbody=="Lake Superior")

choiceData.cov<-choiceData%>%
  mutate(Date=mdy(Date),
         month=as.character(month(Date, label=T)),
         weekday=wday(Date, label=T),
         weekendHoliday=ifelse(weekday%in%c("Sun","Sat") | Date%in%holidays, 1, 0),
         coastal_greatLake=ifelse(PrimaryCounty%in%greatLake$county, 1, 0),
         coastal_river=ifelse(PrimaryCounty%in%river$county, 1, 0),
         coastal_lakeMichigan=ifelse(PrimaryCounty%in%lakeMichigan$county, 1, 0),
         coastal_lakeSuperior=ifelse(PrimaryCounty%in%lakeSuperior$county, 1, 0),
         age=2022-YearOfBirth)%>%
  select(SurveyID, age, Gender, PrimaryCounty, Date, fishedEver, FishedYN, WaterFished, County, Hours:waterbodyType,
         coastal_greatLake, coastal_river, coastal_lakeMichigan, coastal_lakeSuperior, month, weekday, weekendHoliday)%>%
  mutate(age.sc=scale(age, center=T, scale=T),
         choice.partic=ifelse(FishedYN==0, "noFish", "fish"),
         choice.waterbody=ifelse(FishedYN==0, "noFish",
                       ifelse(waterbodyType=="river", "river",
                              ifelse(waterbodyType=="greatLake", "greatLake",
                                     ifelse(waterbodyType=="inlandLake", "inlandLake","error")))),
         choice=ifelse(fishedEver==0, "neverFished",
                       ifelse(FishedYN==0, "notToday",
                              ifelse(waterbodyType=="river","river",
                                     ifelse(waterbodyType=="greatLake", "greatLake",
                                            ifelse(waterbodyType=="inlandLake", "inlandLake", "error"))))),
         choice.specificLake=ifelse(FishedYN==0, "noFish", 
                                    ifelse(waterbodyType=="river","river",
                                           ifelse(waterbodyType=="inlandLake", "inlandLake",
                                                  ifelse(WaterFished=="Lake Michigan", "lakeMichigan",
                                                         ifelse(WaterFished=="Lake Superior", "lakeSuperior", "error"))))),
         age_weekend=age.sc*weekendHoliday,
         female=ifelse(Gender=="FEMALE", 1, 0))%>%
  mutate(shoulder=ifelse(month%in%c("Oct","Nov","Mar","Apr"), 1, 0),
         peakSummer=ifelse(month%in%c("May","Jun","Jul"), 1, 0))


choiceData.dum<-dummy_cols(choiceData.cov, select_columns="month",
                           remove_first_dummy=F,
                           remove_selected_columns=T)%>%
  filter(!is.na(female))

```

Troubleshooting weird negative relationship between living in a coastal county and fishing in the Great Lakes. Dividing Superior from Michigan may help, but let's visualize:

```{r}
summary.nonspecific<-choiceData.cov%>%
  group_by(choice.specificLake, female)%>%
  summarize(n=n())
summary.nonspecific
```
in aggregate, it does look like coastal county anglers are more likely to fish their great lake, especially Superior. (I had it backwards, huh)
Data with only anglers, for comparison (no non-participants)
```{r}

choiceData.cov.f<-choiceData%>%
  mutate(Date=mdy(Date),
         month=as.character(month(Date, label=T)),
         weekday=wday(Date, label=T),
         weekendHoliday=ifelse(weekday%in%c("Sun","Sat") | Date%in%holidays, 1, 0),
         coastal_greatLake=ifelse(PrimaryCounty%in%greatLake$county, 1, 0),
         coastal_river=ifelse(PrimaryCounty%in%river$county, 1, 0),
         age=2022-YearOfBirth)%>%
  select(SurveyID, age, Gender, PrimaryCounty, Date, fishedEver, FishedYN, WaterFished, County, Hours:waterbodyType,
         coastal_greatLake, coastal_river, month, weekday, weekendHoliday)%>%
  mutate(age.sc=scale(age, center=T, scale=T),
         choice.partic=ifelse(FishedYN==0, "noFish", "fish"),
         choice.waterbody=ifelse(FishedYN==0, "noFish",
                       ifelse(waterbodyType=="river", "river",
                              ifelse(waterbodyType=="greatLake", "greatLake",
                                     ifelse(waterbodyType=="inlandLake", "inlandLake","error")))),
         age_weekend=age.sc*weekendHoliday)%>%
  filter(fishedEver==1)%>%
  mutate(shoulder=ifelse(month%in%c("Oct","Nov","Mar","Apr"), 1, 0),
         peakSummer=ifelse(month%in%c("May","Jun","Jul"), 1, 0))



choiceData.dum.f<-dummy_cols(choiceData.cov.f, select_columns="month",
                           remove_first_dummy=F,
                           remove_selected_columns=T)
```
Jumping straight to  nested logit didn't work, shocking no one but myself. 

Start simpler with logit: fishing vs not

```{r}

### Initialise code
apollo_initialise()

### Set core controls
apollo_control = list(
  modelName       = "logit_participation",
  modelDescr      = "Simple logit model--participation",
  indivID         = "SurveyID", 
  outputDirectory = "output_logit_participation"
)

# ################################################################# #
#### LOAD DATA AND APPLY ANY TRANSFORMATIONS                     ####
# ################################################################# #

### Loading data from package
### if data is to be loaded from a file (e.g. called data.csv), 
### the code would be: database = read.csv("data.csv",header=TRUE)
database = choiceData.dum
### for data dictionary, use ?apollo_modeChoiceData


# ################################################################# #
#### DEFINE MODEL PARAMETERS                                     ####
# ################################################################# #

### Vector of parameters, including any that are kept fixed in estimation
apollo_beta=c(asc_noFish=1,
              asc_fish=0,
               b_weekend_holiday=0,
              # b_shoulder=0,
              # b_peakSummer=0,
               b_age=0,
               b_female=0,
               b_age_weekend = 0,
                b_month_Feb=0,
                b_month_Mar=0,
                b_month_Apr=0,
                b_month_May=0,
                b_month_Jun=0,
                b_month_Jul=0,
                b_month_Aug=0,
                b_month_Sep=0,
                b_month_Oct=0,
                # no November data
                b_month_Dec=0
              )

### Vector with names (in quotes) of parameters to be kept fixed at their starting value in apollo_beta, use apollo_beta_fixed = c() if none
apollo_fixed = c("asc_noFish")

# ################################################################# #
#### GROUP AND VALIDATE INPUTS                                   ####
# ################################################################# #

apollo_inputs = apollo_validateInputs()

# ################################################################# #
#### DEFINE MODEL AND LIKELIHOOD FUNCTION                        ####
# ################################################################# #

apollo_probabilities=function(apollo_beta, apollo_inputs, functionality="estimate"){

  ### Attach inputs and detach after function exit
  apollo_attach(apollo_beta, apollo_inputs)
  on.exit(apollo_detach(apollo_beta, apollo_inputs))

  ### Create list of probabilities P
  P = list()
  
  ### List of utilities: these must use the same names as in mnl_settings, order is irrelevant
  V = list()
  V[["noFish"]]  = asc_noFish
  V[["fish"]]  = asc_fish + b_weekend_holiday * weekendHoliday + b_month_Feb * month_Feb + 
    b_month_Mar * month_Mar + b_month_Apr * month_Apr + b_month_May * month_May +
    b_month_Jun * month_Jun + b_month_Jul * month_Jul + b_month_Aug * month_Aug +
    b_month_Sep * month_Sep + b_month_Oct * month_Oct + b_month_Dec * month_Dec +
    b_age * age.sc + b_age_weekend * age_weekend + b_female * female


  
  ### Define settings for MNL model component
  mnl_settings = list(
    alternatives  = c(noFish="noFish", fish="fish"), 
    choiceVar     = choice.partic,
    utilities     = V
  )
  
  ### Compute probabilities using MNL model
  P[["model"]] = apollo_mnl(mnl_settings, functionality)

  ### Take product across observation for same individual
  P = apollo_panelProd(P, apollo_inputs, functionality)

  ### Prepare and return outputs of function
  P = apollo_prepareProb(P, apollo_inputs, functionality)
  return(P)
}

# ################################################################# #
#### MODEL ESTIMATION                                            ####
# ################################################################# #

model = apollo_estimate(apollo_beta, apollo_fixed, apollo_probabilities, apollo_inputs)

# ################################################################# #
#### MODEL OUTPUTS                                               ####
# ################################################################# #

# ----------------------------------------------------------------- #
#---- FORMATTED OUTPUT (TO SCREEN)                               ----
# ----------------------------------------------------------------- #

apollo_modelOutput(model)

# ----------------------------------------------------------------- #
#---- FORMATTED OUTPUT (TO FILE, using model name)               ----
# ----------------------------------------------------------------- #

apollo_saveOutput(model)


```

Instead of nested logit (tried below--problems with nesting parameter), trying normal MNL with covariates. Got weird result for effects of living in a coastal county--trying again with specific coasts
If that works out, try MMNL

```{r}
# ################################################################# #
#### LOAD LIBRARY AND DEFINE CORE SETTINGS                       ####
# ################################################################# #


### Initialise code
apollo_initialise()

### Set core controls
apollo_control = list(
  modelName       = "MNL_participate_waterbodyType",
  modelDescr      = "MNL participation and waterbody type",
  indivID         = "SurveyID", 
  outputDirectory = "output_mnl_fish_waterbody"
)

# ################################################################# #
#### LOAD DATA AND APPLY ANY TRANSFORMATIONS                     ####
# ################################################################# #

### Loading data from package
### if data is to be loaded from a file (e.g. called data.csv), 
### the code would be: database = read.csv("data.csv",header=TRUE)
database = choiceData.dum
### for data dictionary, use ?apollo_modeChoiceData


# ################################################################# #
#### DEFINE MODEL PARAMETERS                                     ####
# ################################################################# #

### Vector of parameters, including any that are kept fixed in estimation. no observations in november, left out on purpose. 
apollo_beta=c( asc_noFish=0,
               asc_fish=0,
               asc_lakeMichigan=0,
               asc_lakeSuperior=0,
               asc_river=0,
              b_weekend_holiday=0,
              b_age=0,
              b_female=0,
              b_weekend_age=0,
                b_month_Feb=0,
                b_month_Mar=0,
                b_month_Apr=0,
                b_month_May=0,
                b_month_Jun=0,
                b_month_Jul=0,
                b_month_Aug=0,
                b_month_Sep=0,
                b_month_Oct=0,
                # no November data
                b_month_Dec=0,
              b_coastal_lakeMichigan=0,
              b_coastal_lakeSuperior=0,
              b_coastal_river=0

              )

### Vector with names (in quotes) of parameters to be kept fixed at their starting value in apollo_beta, use apollo_beta_fixed = c() if none
apollo_fixed = c("asc_noFish")

# ################################################################# #
#### GROUP AND VALIDATE INPUTS                                   ####
# ################################################################# #

apollo_inputs = apollo_validateInputs()

# ################################################################# #
#### DEFINE MODEL AND LIKELIHOOD FUNCTION                        ####
# ################################################################# #

apollo_probabilities=function(apollo_beta, apollo_inputs, functionality="estimate"){

  ### Attach inputs and detach after function exit
  apollo_attach(apollo_beta, apollo_inputs)
  on.exit(apollo_detach(apollo_beta, apollo_inputs))

  ### Create list of probabilities P
  P = list()
  
  ### create alternative specific constants
  
  asc_fish_value = asc_fish + b_weekend_holiday * weekendHoliday + b_month_Feb * month_Feb + 
    b_month_Mar * month_Mar + b_month_Apr * month_Apr + b_month_May * month_May +
    b_month_Jun * month_Jun + b_month_Jul * month_Jul + b_month_Aug * month_Aug +
    b_month_Sep * month_Sep + b_month_Oct * month_Oct + b_month_Dec * month_Dec +
    b_age * age.sc + b_weekend_age * age_weekend + b_female * female
  
  asc_lakeMichigan_value = asc_lakeMichigan +b_coastal_lakeMichigan * coastal_lakeMichigan
  asc_lakeSuperior_value = asc_lakeSuperior + b_coastal_lakeSuperior * coastal_lakeSuperior
  
  ### List of utilities: these must use the same names as in mnl_settings, order is irrelevant

  V = list()
  V[["noFish"]] = asc_noFish 
  V[["river"]] = asc_fish_value + asc_river + b_coastal_river * coastal_river
  V[["lakeMichigan"]] =  asc_fish_value + asc_lakeMichigan_value 
  V[["lakeSuperior"]] = asc_fish_value + asc_lakeSuperior_value 
  V[["inlandLake"]] = asc_fish_value 

 ### Define settings for MNL model component
  mnl_settings = list(
    alternatives = c(noFish="noFish", river="river", lakeMichigan="lakeMichigan", lakeSuperior="lakeSuperior", inlandLake="inlandLake"),
    choiceVar    = choice.specificLake,
    utilities    = V
  )
  
  ### Compute probabilities using MNL model
  P[["model"]] = apollo_mnl(mnl_settings, functionality)
  
  ### Take product across observation for same individual
  P = apollo_panelProd(P, apollo_inputs, functionality)
  
  ### Prepare and return outputs of function
  P = apollo_prepareProb(P, apollo_inputs, functionality)
  return(P)
}

# ################################################################# #
#### MODEL ESTIMATION                                            ####
# ################################################################# #

model = apollo_estimate(apollo_beta, apollo_fixed, apollo_probabilities, apollo_inputs)

# ################################################################# #
#### MODEL OUTPUTS                                               ####
# ################################################################# #

# ----------------------------------------------------------------- #
#---- FORMATTED OUTPUT (TO SCREEN)                               ----
# ----------------------------------------------------------------- #

apollo_modelOutput(model)

# ----------------------------------------------------------------- #
#---- FORMATTED OUTPUT (TO FILE, using model name)               ----
# ----------------------------------------------------------------- #

apollo_saveOutput(model)

# ################################################################# #
##### POST-PROCESSING                                            ####
# ################################################################# #

### Print outputs of additional diagnostics to new output file (remember to close file writing when complete)
apollo_sink()
```
Aha, adding ASCs for great lakes and rivers fixed the weird negative parameter value for coastal_lakeMichigan


Predicting fishing effort using this model.

I'm going to predict probabilities across a full year for an 'average' angler

but first, predictions for the data I have

```{r}
model=apollo_loadModel("MNL_participate_waterbodyType")

predictions_base = apollo_prediction(model, apollo_probabilities, apollo_inputs)
```


Plan: 
- Make predictions of probs for a year assuming identical anglers
- mess with ages--how much variation is there?
- Work on (roughly) incorporating into sim model--how do I want to handle 'bins' for non-fishing, non-inland lake options? Assume that everyone in the sim made the choice to fish? Or include non-participation? (compare fits of full data vs. only participants)
- Incorporate heterogeneity--add random component, look for correlations among (random) coefficients (regression approach)
- Add individual heterogeneity into the model
- Move on to lake choice--for those choosing inland lakes, adapt DCE estimates (run some test for how to deal with distance--straight line at some (slow) speed approaching actual travel distance? OR, place anglers at population centers, for which I have estimated distances to each lake? Start simple with random placement and straight line distance at slow speed. Then adapt angler population density from there)


```{r}
predict_data<-data.frame(date=seq(mdy("1/1/2021"),mdy("12/31/2021"), by="days"))

# eventually include a range of ages (repeated sims)
pred_data_mean<-predict_data%>%
  mutate(SurveyID=rep(1),
         choice.specificLake=rep(NA),
         day=wday(date, label=T),
         month=month(date, label=T))%>%
  dummy_cols(select_columns="month", remove_first_dummy=F, remove_selected_columns=T)%>%
  mutate(weekendHoliday=ifelse(day%in%c("Sat","Sun") | date%in%holidays, 1, 0),
         age.sc=rep(0),
         age_weekend=weekendHoliday*age.sc,
         coastal_lakeMichigan=0,
         coastal_lakeSuperior=0,
         coastal_river=0,
         female=0)


database<-pred_data_mean
apollo_inputs=apollo_validateInputs()
predictions_new=apollo_prediction(model, apollo_probabilities, apollo_inputs)
```

Plot probabilities over time

```{r}
predictions_pivot<-predictions_new%>%
  pivot_longer(cols=c(noFish:inlandLake), names_to="choice",values_to="probChoice")

ggplot(predictions_pivot)+
  geom_line(aes(x=Observation, y=probChoice, color=choice))

# ok, the expected seasonal pattern is there. Filter out prob noFish to look at different waterbodies

predictions_fished<-predictions_pivot%>%
  filter(choice!="noFish")

ggplot(predictions_fished)+
  geom_line(aes(x=Observation, y=probChoice, color=choice))

```
neat, but are those 0 probabilities for great lakes? 

Make a prediction df for different ages and coastal/noncoastal counties

```{r}
anglerInland<-data.frame(SurveyID=c(1:3), age.sc=c(-1, 0, 1), coastal_lakeMichigan=0, coastal_lakeSuperior=0, coastal_river=0)
anglerRiver<-data.frame(SurveyID=c(4:6), age.sc=c(-1, 0, 1), coastal_lakeMichigan=0, coastal_lakeSuperior=0, coastal_river=1)
anglerMich<-data.frame(SurveyID=c(7:9), age.sc=c(-1, 0, 1), coastal_lakeMichigan=1, coastal_lakeSuperior=0, coastal_river=0)
anglerSup<-data.frame(SurveyID=c(10:12), age.sc=c(-1, 0, 1), coastal_lakeMichigan=0, coastal_lakeSuperior=1, coastal_river=0)

anglerSim.1<-rbind.data.frame(anglerInland, anglerRiver, anglerMich, anglerSup)%>%
  mutate(female=1)

anglerSim.2<-rbind.data.frame(anglerInland, anglerRiver, anglerMich, anglerSup)%>%
  mutate(female=0)

num.row<-nrow(anglerSim.1)+nrow(anglerSim.2)

anglerSim<-rbind.data.frame(anglerSim.1, anglerSim.2)%>%
  mutate(SurveyID=seq(1:num.row))

  # then repeat each row 365 times

pred_data<-anglerSim%>%
  slice(rep(1:n(), each=365))%>%
  mutate(date=rep(seq(mdy("1/1/2021"),mdy("12/31/2021"), by="days"), length(unique(SurveyID))),
         choice.specificLake=rep(NA),
         day=wday(date, label=T),
         month=month(date, label=T))%>%
  dummy_cols(select_columns="month", remove_first_dummy=F, remove_selected_columns=F)%>%
  mutate(weekendHoliday=ifelse(day%in%c("Sat","Sun") | date%in%holidays, 1, 0),
         month_Mar=ifelse(month%in%c("Mar","Nov"), 1, 0),
         age_weekend=weekendHoliday*age.sc)

         # adding dummy indicator for March in November (no observations--other shoulder season

database=pred_data
apollo_inputs=apollo_validateInputs()
predictions=apollo_prediction(model, apollo_probabilities, apollo_inputs)

predictions_pivot<-predictions%>%
  mutate(age.sc=pred_data$age.sc,
         coastal_lakeMichigan=pred_data$coastal_lakeMichigan,
         coastal_lakeSuperior=pred_data$coastal_lakeSuperior,
         coastal_river=pred_data$coastal_river,
         female=pred_data$female)%>%
  pivot_longer(cols=c(noFish:inlandLake), names_to="choice", values_to="choiceProb")%>%
  mutate(choice=as.factor(choice),
         ID=as.factor(ID),
         age=ifelse(age.sc==-1, "younger",
                    ifelse(age.sc==0, "average", "older")),
         residence=ifelse(coastal_lakeMichigan==1, "lakeMichigan",
                          ifelse(coastal_lakeSuperior==1, "lakeSuperior",
                                 ifelse(coastal_river==1, "river", "inland"))),
         gender=ifelse(female==1, "female","male"))%>%
  filter(choice!="noFish")

ggplot(predictions_pivot)+
  geom_line(aes(x=Observation, y=choiceProb, color=residence, linetype=age))+
  scale_color_manual(values=brewer.pal(n=4, "Set1"))+
  facet_grid(choice~gender)+
  theme_bw()

# wide format predictions with covariates
predictions_cov<-predictions%>%
  mutate(gender=ifelse(pred_data$female==1, "female", "male"),
         age.sc=pred_data$age.sc,
         coastal_lakeMichigan=pred_data$coastal_lakeMichigan,
         coastal_lakeSuperior=pred_data$coastal_lakeSuperior,
         coastal_river=pred_data$coastal_river)
```
Hard to parse age differences
```{r}
ggplot(predictions_pivot)+
  geom_line(aes(x=Observation, y=choiceProb, color=age, linetype=residence))+
  scale_color_manual(values=brewer.pal(n=4, "Set1"))+
  facet_grid(choice~gender)+
  theme_bw()


```
yeah, age primarily drives distribution through the week (which makes sense; I didn't include it in waterbody type choice)

So now see what that looks like in terms of fishign days/year. Say we have 800,000 licenses active in a year. 

uhhh I can't find any of the DNR's social science reports, what? Ideally, I would like cross tabs of license sales by county, with breakdowns of age and gender.

For now, I'll just use county populations from the Census, assuming similar participation rates across counties and 50/50 gender split. At some point, I need to get cross tabs of license sales data. (by county, age, gender) (gender more important than age if there are limitations)

Some estimates of fishing effort on inland lakes and Great lakes:
- From Holly Embke's paper estimating WI fishing effort--from plot of trends in directed fishing effort, eyeballed mean, multiplied by number of species, (gave hours/ha) multiplied by total surface area (305693) divided by 4 (assuming 4 hour fishing trips)
3,363,612 fishing trips on inland lakes

- From creel survey reports, also assuming 4 hour trips: 155,581 trips on Lake Michigan and 80,351 on Superior in 2022

- nothing on rivers, so leaving them out
```{r}
total.obs<-3363612+155581+80351

inland<-3363612/total.obs

great<-(155581+80351)/total
inland
great
```
model outputs (for fishing inland and great lakes) are about 98% inland, 2% great lakes. 

```{r}
coastalCounties<-read_csv(here::here("data","coastalCounties.csv"))

superior<-filter(coastalCounties, waterbody=="Lake Superior")
michigan<-filter(coastalCounties, waterbody=="Lake Michigan")
river<-filter(coastalCounties, waterbody%in%c("Mississippi River","Wisconsin River"))

# narrowing down predictions--dropping variation in age, assuming mean age for all predictions
predictions_meanAge<-predictions_cov%>%
  cbind.data.frame(dplyr::select(pred_data, month, month_Jan:month_Dec, weekendHoliday))%>%
  filter(age.sc==0)

# pulling out distinct predictions (i.e. by month and weekday type) for join to effort estimate data
predictions_distinct<-predictions_meanAge%>%
  select(ID, noFish:weekendHoliday)%>%
  distinct()

yearDates<-data.frame(date=seq(mdy("1/1/2021"),mdy("12/31/2021"), by="days"))%>%
  mutate(month=month(date, label=T),
         wday=wday(date, label=T),
         weekendHoliday=ifelse(wday%in%c("Sat","Sun") | date%in%holidays, 1, 0))%>%
  group_by(month, weekendHoliday)%>%
  summarize(n=n())

gender<-rep(c("female","male"), 4)



pop<-read_csv(here::here("data","county.pop.csv"))%>%
  mutate(county=str_split_fixed(`Label (Grouping)`, " County,", 2)[,1])%>%
  rename("population"=`Total population`)%>%
  select(county, population)%>%
  mutate(coastal_lakeSuperior=ifelse(county%in%superior$county, 1, 0),
         coastal_lakeMichigan=ifelse(county%in%michigan$county, 1, 0),
         coastal_river=ifelse(county%in%river$county, 1, 0))%>%
  group_by(coastal_lakeSuperior, coastal_lakeMichigan, coastal_river)%>%
  # (for now) assuming 25% of county population has a license
  summarize(population=sum(population)*0.25)%>%
  ungroup()%>%
  slice(rep(1:n(), each=2))%>%
  mutate(gender=rep(c("female","male"), 4),
         # population of anglers; i.e. rough guess that 25% of population hold license
         population=round((population/2)*0.25),
         age=rep(0),
         stratumID=seq(1:8))%>%
  # now slice so that each row (type of individual) is repeated for a year
    slice(rep(1:n(), each=24))%>%
  mutate(month=rep(yearDates$month, 8),
         weekendHoliday=rep(yearDates$weekendHoliday, 8),
         nDays=rep(yearDates$n, 8))%>%
  dummy_cols(select_columns="month", remove_first_dummy=F, remove_selected_columns=F)%>%
  # no November observations--set to March to account for shoulder season
  mutate(month_March=ifelse(month%in%c("Nov","Mar"),1,0))%>%

  # now join to the choice probability predictions by month, weekendHoliday, coastal vars, and gender
  left_join(predictions_distinct[,c("month","weekendHoliday","coastal_lakeMichigan",
                                    "coastal_lakeSuperior","coastal_river","gender",
                                    "noFish","river","lakeMichigan","lakeSuperior",
                                    "inlandLake")], by=c("month","weekendHoliday", 
                                    "coastal_lakeMichigan",
                                    "coastal_lakeSuperior","coastal_river","gender"))%>%
  #Then multiply prob by stratum population, multiply that by days in the stratum, and add across strata to get annual estimates
  mutate(nLatentDays=population*nDays,
         nNoFishingDays=nLatentDays*noFish,
         nRiverDays=nLatentDays*river,
         nLakeMichiganDays=nLatentDays*lakeMichigan,
         nLakeSuperiorDays=nLatentDays*lakeSuperior,
         nInlandDays=nLatentDays*inlandLake)

nLatent<-sum(pop$nLatentDays)
nNoFish<-sum(pop$nNoFishingDays)
nLakeMichigan<-sum(pop$nLakeMichiganDays)
nLakeSuperior<-sum(pop$nLakeSuperiorDays)
nRiver<-sum(pop$nRiverDays)
nInland<-sum(pop$nInlandDays)

total<-nLakeMichigan+nLakeSuperior+nInland
nInland/total

# that's a lot of River effort predicted (734886 angler trips/year)
```
Neat! Proportions predicted by the model are pretty similar (0.93 inland from model vs 0.95 inland from observations). Numbers are a bit off (lower in absolute terms), but proportion inland  vs great lakes is pretty similar. Predictions for the two great lakes are much more off--very, very low for Superior (0.06 of the observed effort), also low for Michigan (0.5 observed effort)

So I'll want to revisit this with the improved license data. 

suggestion to instead try mixed logit for panel data https://www.statalist.org/forums/forum/general-stata-discussion/general/1569500-nested-logit-with-panel-data 


    